---
title: "Timeliness of UKHSA datafeeds for mandatory reporting of HCAIs"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(magrittr)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(reactable)

```

```{r}
# Load data

monitor_results <- read_csv(
  "./ukhsa_datafeeds_timeliness_anon.csv",
  col_types = cols(
    monitor_datetime = col_datetime(),
    monitor_point_name = col_character(),
    monitor_point_subname = col_character(),
    stat_name = col_character(),
    stat_type = col_character(),
    columnname = col_character(),
    groupby_colname = col_character(),
    groupby_value = col_character(),
    value = col_character(),
    value_type = col_character(),
    run_date = col_date()
  )
)

```


```{r}
cat("Latest monitor date: ", format(max(monitor_results$monitor_datetime)))

audit_dates <-
  monitor_results %>%
  mutate(audit_dates = substring(monitor_point_subname, nchar(monitor_point_subname) - 7)) %>%
  distinct(audit_dates) %>%
  arrange(audit_dates) %>%
  pull()

cat("audit_dates:", audit_dates)
tibble(audit_dates = audit_dates,
       followup_days = as.Date(max(monitor_results$monitor_datetime)) - lubridate::ymd(audit_dates))

```

```{r}
stat_line_plot <- function(df, stat, audit_date){
  g <- 
    ggplot(data = df %>% filter(stat_name == stat),
           aes(x = run_date, y = num_value)) +
    scale_x_date(
      limits = c(lubridate::ymd(audit_date), NA),
      breaks = scales::breaks_pretty(),
      labels = scales::label_date_short(sep = " "),
    ) +
    labs(x = "Monitor date",
                  y = stat,
                  title = NULL) +
    theme_bw() +
    geom_line()
  
  g
}

stat_threshold_table <- function(df, audit_date){
  df %>%
    group_by(stat_name) %>%
    arrange(run_date) %>%
    mutate(total = last(num_value),
           perc = 100*num_value/total,
           dayspassed = as.integer(run_date - lubridate::ymd(audit_date)),
           days50 = min(dayspassed[perc >= 50]),
           days90 = min(dayspassed[perc >= 90]),
           days95 = min(dayspassed[perc >= 95]),
           days99 = min(dayspassed[perc >= 99]),
           ) %>%
    slice(1) %>%
    select(stat_name, total, days50, days90, days95, days99)
}

```

```{r}
# initialise htmlwidgets otherwise they won't render
# this chunk can be placed anywhere
dummy_df <- data.frame(a = as.Date("2023-01-01"), b = "item", c = 1)

reactable(dummy_df, 
          height = 0,
          bordered = FALSE)
```

## SGSS {.tabset}

Organisms included: o.Organism_Species_Name IN ('ESCHERICHIA COLI','STREPTOCOCCUS PNEUMONIAE','STAPHYLOCOCCUS AUREUS','KLEBSIELLA OXYTOCA','KLEBSIELLA PNEUMONIAE','KLEBSIELLA PNEUMONIAE SUBSP OZENAE','KLEBSIELLA PNEUMONIAE SUBSP PNEUMONIAE')
		OR o.Organism_Genus_Name IN ('ENTEROCOCCUS','ACINETOBACTER','PSEUDOMONAS')

```{r}
sgss_by_organism <- function(monitor_results, record_type, audit_date){
    monitor_results %>%
    filter(monitor_point_name == "sgss_specdates") %>%
    filter(monitor_point_subname == paste0(record_type, "_", audit_date)) %>%
    filter(stat_name == "Specimens per organism") %>%
    filter(run_date >= lubridate::ymd(audit_date)) %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(run_date, stat_name, groupby_colname, organism = groupby_value, num_value) %>%
    arrange(run_date)
}

sgss_data <- function(monitor_results, record_type, audit_date, stats){
    monitor_results %>%
    filter(monitor_point_name == "sgss_specdates") %>%
    filter(monitor_point_subname == paste0(record_type, "_", audit_date)) %>%
    filter(stat_name %in% stats) %>%
    filter(run_date >= lubridate::ymd(audit_date)) %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(run_date, stat_name, groupby_colname, groupby_value, num_value) %>%
    arrange(run_date)
}


sgss_by_organism_plot <- function(df, audit_date){
  legend_order <-
    df %>%
    group_by(organism) %>%
    arrange(desc(run_date)) %>%
    slice(1) %>%
    arrange(desc(num_value)) %>%
    pull(organism)

  df$organism <- factor(df$organism, levels = legend_order)

  minx <- lubridate::ymd(audit_date)
  
  g <- ggplot(data = df,
              aes(x = run_date, y = num_value, group = organism)) +
    scale_x_date(
      limits = c(minx, NA),
      breaks = scales::breaks_pretty(),
      labels = scales::label_date_short(sep = " "),
    ) +
    labs(x = "Monitor date",
         y = "No. of records") +
    theme_bw() +
    guides(color = guide_legend(title = "Organism")) +
    theme(
      legend.position = "inside",
      legend.justification.inside = c(1, 0),
      legend.position.inside = c(0.92, 0.2),
      legend.background = element_rect(colour = "grey")
    ) +
    geom_line(aes(color = organism))
  
  g
}

sgss_by_organism_thresholds <- function(df, audit_date){
  df %>%
    group_by(organism) %>%
    arrange(run_date) %>%
    mutate(total = last(num_value),
           perc = 100*num_value/total,
           dayspassed = as.integer(run_date - lubridate::ymd(audit_date)),
           days50 = min(dayspassed[perc >= 50]),
           days90 = min(dayspassed[perc >= 90]),
           days95 = min(dayspassed[perc >= 95]),
           days99 = min(dayspassed[perc >= 99]),
           ) %>%
    slice(1) %>%
    select(stat_name, organism, total, days50, days90, days95, days99)
}

```



### CDR {.tabset}

#### By specimen {.tabset}


```{r, results='asis'}

for (audit_date in audit_dates) {
  cat("\n#####", audit_date, "\n\n")
  
  df <- sgss_by_organism(monitor_results, "cdr_opiedate", audit_date)
  
  sgss_by_organism_plot(df, audit_date) %>%
    print()
  
  cat("\n")
  
  thresholds <- sgss_by_organism_thresholds(df, audit_date) %>%
    reactable()
  
  # NOTE: a regular print() doesn't render the widget
  cat(knitr::knit_print(thresholds))
  
  cat("\n")
}

```


#### Labs {.tabset}

```{r, results='asis'}

for (audit_date in audit_dates) {
  cat("\n#####", audit_date, "\n\n")
  
  labsdf <- sgss_data(monitor_results, "cdr_opiedate", audit_date, "No. of Labs")
  
  stat_line_plot(labsdf, stat = "No. of Labs", audit_date) %>%
    print()
  
  cat("\n")
  
  thresholds <- stat_threshold_table(labsdf, audit_date) %>%
    reactable()
  
  # NOTE: a regular print() doesn't render the widget
  cat(knitr::knit_print(thresholds))
  
  cat("\n")
}

```


### AMR {.tabset}


#### By specimen {.tabset}

```{r, results='asis'}

for (audit_date in audit_dates) {
  cat("\n#####", audit_date, "\n\n")
  
  df <- sgss_by_organism(monitor_results, "amr_specdate", audit_date)
  
  sgss_by_organism_plot(df, audit_date) %>%
    print()
  
  cat("\n")
  
  thresholds <- sgss_by_organism_thresholds(df, audit_date) %>%
    reactable()
  
  # NOTE: a regular print() doesn't render the widget
  cat(knitr::knit_print(thresholds))
  
  cat("\n")
}

```


#### Labs {.tabset}

```{r, results='asis'}

for (audit_date in audit_dates) {
  cat("\n#####", audit_date, "\n\n")
  
  labsdf <- sgss_data(monitor_results, "amr_specdate", audit_date, "No. of Labs")
  
  stat_line_plot(labsdf, stat = "No. of Labs", audit_date) %>%
    print()
  
  cat("\n")
  
  thresholds <- stat_threshold_table(labsdf, audit_date) %>%
    reactable()
  
  # NOTE: a regular print() doesn't render the widget
  cat(knitr::knit_print(thresholds))
  
  cat("\n")
}

```


## SUS+ {.tabset}

```{r}
sus_apc_data <- function(monitor_results, audit_date){
  monitor_results %>%
    filter(monitor_point_name == "sus_dates") %>%
    filter(monitor_point_subname == paste0("apc_adm_date_", audit_date)) %>%
    filter(stat_name %in% c("Total admissions", "Total discharged", "Total with diagcodes", "No. of Providers")) %>%
    filter(run_date >= lubridate::ymd(audit_date)) %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(run_date, stat_name, num_value)
}

sus_ecds_data <- function(monitor_results, audit_date){
  monitor_results %>%
    filter(monitor_point_name == "sus_dates") %>%
    filter(monitor_point_subname == paste0("ecds_arr_date_", audit_date)) %>%
    filter(stat_name %in% c("Total arrivals", "No. of Providers")) %>%
    filter(run_date >= lubridate::ymd(audit_date)) %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(run_date, stat_name, num_value)
}


sus_apc_episodes_plot <- function(df, audit_date){
  plotdata <- df %>%
    filter(stat_name %in% c("Total admissions", "Total discharged", "Total with diagcodes")) %>%
    mutate(record_type = factor(stat_name, 
                                levels = c("Total admissions", 
                                           "Total discharged", 
                                           "Total with diagcodes"),
                                labels = c("Admitted, no dischargedate/diagcodes", 
                                           "Discharged, no diagcodes", 
                                           "Discharged, with diagcodes"))
           )

  g <- ggplot(data = plotdata, 
              aes(x = run_date, y = num_value, fill = record_type)) +
  		scale_x_date(
  		  limits = c(lubridate::ymd(audit_date), NA),
  			breaks = scales::breaks_pretty(),
  			labels = scales::label_date_short(sep = " "),
  		) +
  		labs(
  			x = "Monitor date",
  			y = "No. of records",
  			title = NULL
  		) +
    theme_bw() +
    theme(
      legend.position = "inside",
      legend.justification.inside = c(1, 0),
      legend.position.inside = c(0.92, 0.3),
      legend.background = element_rect(colour = "grey")
    ) +
    guides(fill = guide_legend(title = "Record type", direction = "vertical")) +
    scale_fill_viridis_d() +
    geom_area(position = "identity")
  
  g
}

```

### APC totals {.tabset} 

#### Records {.tabset}

```{r, results='asis'}

for (audit_date in audit_dates) {
  cat("\n#####", audit_date, "\n\n")
  
  df <- sus_apc_data(monitor_results, audit_date)
  
  sus_apc_episodes_plot(df, audit_date) %>%
    print()
  
  cat("\n")
    
  thresholds <- stat_threshold_table(df %>%
                                       filter(stat_name != "No. of Providers"),
                                     audit_date) %>%
    reactable()
    
  
  # NOTE: a regular print() doesn't render the widget
  cat(knitr::knit_print(thresholds))
  
  cat("\n")
}

```


#### Providers {.tabset}

```{r, results='asis'}

for (audit_date in audit_dates) {
  cat("\n#####", audit_date, "\n\n")
  
  df <- sus_apc_data(monitor_results, audit_date)
  
  stat_line_plot(df, "No. of Providers", audit_date) %>%
    print()
  
  cat("\n")
  
  thresholds <- stat_threshold_table(df %>%
                                       filter(stat_name == "No. of Providers"),
                                     audit_date) %>%
    reactable()
    
  
  # NOTE: a regular print() doesn't render the widget
  cat(knitr::knit_print(thresholds))
  
  cat("\n")
}

```

### ECDS totals {.tabset}

#### Records {.tabset}

```{r, results='asis'}

for (audit_date in audit_dates) {
  cat("\n#####", audit_date, "\n\n")
  
  df <- sus_ecds_data(monitor_results, audit_date)
  
  stat_line_plot(df, "Total arrivals", audit_date) %>%
    print()
  
  cat("\n")
    
  thresholds <- stat_threshold_table(df %>%
                       filter(stat_name == "Total arrivals"),
                       audit_date) %>%
    reactable()
  
  # NOTE: a regular print() doesn't render the widget
  cat(knitr::knit_print(thresholds))
  
  cat("\n")
}

```

#### Providers {.tabset}

```{r, results='asis'}

for (audit_date in audit_dates) {
  cat("\n#####", audit_date, "\n\n")
  
  df <- sus_ecds_data(monitor_results, audit_date)
  
  stat_line_plot(df, "No. of Providers", audit_date) %>%
    print()
  
  cat("\n")
  
  thresholds <- stat_threshold_table(df %>%
                       filter(stat_name == "No. of Providers"),
                       audit_date) %>%
    reactable()
  
  # NOTE: a regular print() doesn't render the widget
  cat(knitr::knit_print(thresholds))
  
  cat("\n")
}

```



## Summary {.tabset}


### No. of SGSS records on each audit date {.tabset}

#### All organisms

```{r}
summary_results <-
  monitor_results %>%
    filter(monitor_point_name == "sgss_specdates") %>%
    filter(substr(monitor_point_subname, 1, 12) %in% c("cdr_opiedate", "amr_specdate")) %>%
    mutate(source = toupper(substr(monitor_point_subname, 1, 3)),
           audit_date = substr(monitor_point_subname, 14, 21)) %>%
    filter(stat_name == "Specimens per organism") %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    group_by(source, audit_date, run_date) %>%
    summarise(num_value = sum(num_value),
              .groups = "drop") %>%
    select(run_date, source, audit_date, num_value) %>%
    group_by(source, audit_date) %>%
    arrange(run_date) %>%
    mutate(total = last(num_value),
           perc = 100*num_value/total,
           dayspassed = as.integer(run_date - lubridate::ymd(audit_date)),
           days50 = min(dayspassed[perc >= 50]),
           days90 = min(dayspassed[perc >= 90]),
           days95 = min(dayspassed[perc >= 95]),
           days99 = min(dayspassed[perc >= 99]),
           ) %>%
    slice(1) %>%
    select(source, audit_date, total, days50, days90, days95, days99) %>%
    arrange(source, audit_date) 

summary_results %>%
  reactable::reactable(filterable = TRUE)

```

```{r}
summary_results %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(source, " - ", name)) %>%
  ggplot(aes(x = factor(category, levels = paste0(
                          c(
                            rep("CDR", 4), 
                            rep("AMR", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 2))
                        ),
             y = count, fill = source)) +
		labs(
			x = "Percentile of records received",
			y = "No. of days after specimen date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  theme_bw() +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 3)

```

#### By organism

```{r}
summary_results_by_organism <-
  monitor_results %>%
    filter(monitor_point_name == "sgss_specdates") %>%
    filter(substr(monitor_point_subname, 1, 12) %in% c("cdr_opiedate", "amr_specdate")) %>%
    mutate(source = toupper(substr(monitor_point_subname, 1, 3)),
           audit_date = substr(monitor_point_subname, 14, 21)) %>%
    filter(stat_name == "Specimens per organism") %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(run_date, source, audit_date, organism = groupby_value, num_value) %>%
    group_by(source, organism, audit_date) %>%
    arrange(run_date) %>%
    mutate(total = last(num_value),
           perc = 100*num_value/total,
           dayspassed = as.integer(run_date - lubridate::ymd(audit_date)),
           days50 = min(dayspassed[perc >= 50]),
           days90 = min(dayspassed[perc >= 90]),
           days95 = min(dayspassed[perc >= 95]),
           days99 = min(dayspassed[perc >= 99]),
           ) %>%
    slice(1) %>%
    select(source, audit_date, organism, total, days50, days90, days95, days99) %>%
    arrange(source, organism, audit_date)
```

```{r}
summary_results_by_organism %>%
    reactable::reactable(filterable = TRUE)

```


### No. of labs on each audit date {.tabset}

```{r}
summary_labs <-
  monitor_results %>%
    filter(monitor_point_name == "sgss_specdates") %>%
    filter(substr(monitor_point_subname, 1, 12) %in% c("cdr_opiedate", "amr_specdate")) %>%
    mutate(source = toupper(substr(monitor_point_subname, 1, 3)),
           audit_date = substr(monitor_point_subname, 14, 21)) %>%
    filter(stat_name == "Specimens per lab") %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(run_date, audit_date, source, labname = groupby_value, num_value) %>%
    group_by(source, labname, audit_date) %>%
    arrange(run_date) %>%
    mutate(new_records = num_value - lag(num_value) > 0 | (row_number() == 1 & num_value > 0),
           dayspassed = as.integer(run_date - lubridate::ymd(audit_date))) %>%
    filter(new_records) %>%
    group_by(source, labname, audit_date) %>%
    summarise(lastrecord = max(dayspassed),
              firstrecord = min(dayspassed),
              .groups = "drop")

```

#### Total no. of labs

```{r}
summary_labs %>%
  group_by(source) %>%
  summarise(n_distinct(labname))

```

#### Longest record to arrive


```{r}
summary_labs %>%
  group_by(source) %>%
  summarise(maxfirstrecord = max(firstrecord),
            maxlastrecord = max(lastrecord),
            .groups = "drop")

```

#### First records


```{r}
summary_labs_firstrecord <-
  summary_labs %>%
    group_by(source, audit_date) %>%
    summarise(total = n_distinct(labname),
           days50 = quantile(firstrecord, 0.5, type = 1),
           days90 = quantile(firstrecord, 0.90, type = 1),
           days95 = quantile(firstrecord, 0.95, type = 1),
           days99 = quantile(firstrecord, 0.99, type = 1),
           .groups = "drop"
           )
```


```{r}
summary_labs_firstrecord %>%
  reactable::reactable(filterable = TRUE)
```


```{r}
summary_labs_firstrecord %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(source, "_", name)) %>%
  ggplot(aes(x = factor(category), y = count, fill = source)) +
		labs(
			x = "Percentile of labs sent first record",
			y = "No. of days after specimen date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  theme_bw() +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 3)

```

#### Last records

```{r}
summary_labs_lastrecord <-
  summary_labs %>%
    group_by(source, audit_date) %>%
    summarise(total = n_distinct(labname),
           days50 = quantile(lastrecord, 0.5, type = 1),
           days90 = quantile(lastrecord, 0.90, type = 1),
           days95 = quantile(lastrecord, 0.95, type = 1),
           days99 = quantile(lastrecord, 0.99, type = 1),
           .groups = "drop"
           )
```


```{r}
summary_labs_lastrecord %>%
  reactable::reactable(filterable = TRUE)
```


```{r}
summary_labs_lastrecord %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(source, "_", name)) %>%
  ggplot(aes(x = factor(category), y = count, fill = source)) +
		labs(
			x = "Percentile of labs sent last record",
			y = "No. of days after specimen date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  theme_bw() +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 7)

```

#### First and last records


```{r}
summary_labs_firstlast_record <-
  rbind(tibble(summary_labs_firstrecord, record_position = "first"),
        tibble(summary_labs_lastrecord, record_position = "last"))
```


```{r}
summary_labs_firstlast_record %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(record_position, " - ", source, " - ", name)) %>%
  ggplot(aes(x = factor(category,
                        levels = paste0(
                          c(
                            rep("first - ", 4),
                            rep("last - ", 4)
                            ),
                          c(
                            rep("CDR", 8), 
                            rep("AMR", 8)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        labels = paste0(
                          c(
                            rep("First ", 4),
                            rep("Last ", 4)
                            ),
                          c(rep("CDR record", 8), 
                            rep("AMR record", 8)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        ),
             y = count, fill = source)) +
		labs(
			x = "Percentile of labs submitting records",
			y = "No. of days after specimen date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 7)


```


### No. of SUS+ records on each audit date {.tabset}

#### ECDS disruption

audit dates after 20230824 have majority of records loaded on a single day, and have a spurious date (2023-09-08) with too many records 

```{r}
monitor_results %>%
  filter(monitor_point_name == "sus_dates") %>%
  filter(substr(monitor_point_subname, 1, 13) %in% c("ecds_arr_date")) %>%
  mutate(audit_date = substr(monitor_point_subname, regexpr(pattern = "date_", monitor_point_subname) + 5, 22)) %>%
  filter(stat_name %in% c("Total arrivals")) %>%
  filter(!is.na(value)) %>%
  mutate("num_value" = as.numeric(value)) %>%
  select(audit_date, run_date, subtype = stat_name, num_value) %>%
  group_by(subtype, audit_date) %>%
  arrange(run_date) %>%
reactable::reactable(filterable = TRUE)

```

#### By subtype

(After removing spurious ECDS loading date)

```{r}
summary_results_by_subtype <-
  monitor_results %>%
    filter(monitor_point_name == "sus_dates") %>%
    filter(substr(monitor_point_subname, 1, 13) %in% c("ecds_arr_date", "apc_adm_date_")) %>%
    mutate(source = toupper(substr(monitor_point_subname, 1, regexpr(pattern = "_", monitor_point_subname) - 1)),
           audit_date = substr(monitor_point_subname, regexpr(pattern = "date_", monitor_point_subname) + 5, 22)) %>%
    filter(!(stat_name == "Total arrivals" & run_date == as.Date("2023-09-08"))) %>% # spurious excess loading of ecds data
    filter(stat_name %in% c("Total admissions", "Total discharged", "Total with diagcodes", "Total arrivals")) %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(run_date, source, audit_date, stat_name, num_value) %>%
    group_by(source, stat_name, audit_date) %>%
    arrange(run_date) %>%
    mutate(total = last(num_value),
           perc = 100*num_value/total,
           dayspassed = as.integer(run_date - lubridate::ymd(audit_date)),
           days50 = min(dayspassed[perc >= 50]),
           days90 = min(dayspassed[perc >= 90]),
           days95 = min(dayspassed[perc >= 95]),
           days99 = min(dayspassed[perc >= 99]),
           ) %>%
    slice(1) %>%
    select(source, audit_date, stat_name, total, days50, days90, days95, days99) %>%
    arrange(source, stat_name, audit_date)
```

```{r}
summary_results_by_subtype %>%
    reactable::reactable(filterable = TRUE)

```

```{r}
summary_results_by_subtype %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(stat_name, " - ", name),
         loading = case_when(audit_date %in% c("20230819","20230824","20230829","20230901") ~ "Disrupted",
                                     TRUE ~ "Undisrupted")) %>%
  ggplot(aes(x = factor(category, 
                        levels = paste0(
                          c(
                            rep("Total admissions", 4), 
                            rep("Total discharged", 4), 
                            rep("Total with diagcodes", 4), 
                            rep("Total arrivals", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        labels = paste0(
                          c(
                            rep("Admisions", 4), 
                            rep("Discharged", 4), 
                            rep("Discharged + diagcodes", 4), 
                            rep("ED Arrivals", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        ), 
             y = count, fill = source, color = loading)) +
		labs(
			x = "Percentile of records received",
			y = "No. of days after adm/arr date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  scale_color_manual(values = c("orange", "black")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 4)

```




### No. of providers on each audit date {.tabset}

```{r}
summary_providers <-
  monitor_results %>%
    filter(monitor_point_name == "sus_dates") %>%
    filter(substr(monitor_point_subname, 1, 13) %in% c("ecds_arr_date", "apc_adm_date_")) %>%
    mutate(source = toupper(substr(monitor_point_subname, 1, regexpr(pattern = "_", monitor_point_subname) - 1)),
           audit_date = substr(monitor_point_subname, regexpr(pattern = "date_", monitor_point_subname) + 5, 22)) %>%
    filter(!(stat_name == "Arrivals per Provider" & run_date == as.Date("2023-09-08"))) %>% # spurious excess loading of ecds data
    filter(stat_name %in% c("Admissions per Provider", "Discharged per Provider", "Diags present per Provider", "Arrivals per Provider")) %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(run_date, audit_date, source, stat_name, provider = groupby_value, num_value) %>%
    group_by(source, stat_name, provider, audit_date) %>%
    arrange(run_date) %>%
    mutate(new_records = num_value - lag(num_value) > 0 | (row_number() == 1 & num_value > 0),
           dayspassed = as.integer(run_date - lubridate::ymd(audit_date))) %>%
    filter(new_records) %>%
    group_by(source, stat_name, provider, audit_date) %>%
    summarise(lastrecord = max(dayspassed),
              firstrecord = min(dayspassed),
              .groups = "drop") 

```

#### Total no. of providers

```{r}
summary_providers %>%
  group_by(source) %>%
  summarise(n_distinct(provider))

```

#### Longest record to arrive


```{r}
summary_providers %>%
  group_by(source, stat_name) %>%
  summarise(maxfirstrecord = max(firstrecord),
            maxlastrecord = max(lastrecord),
            .groups = "drop")

```


#### First records

```{r}
summary_providers_firstrecord <-
  summary_providers %>%
    group_by(source, stat_name, audit_date) %>%
    summarise(total = n_distinct(provider),
           days50 = quantile(firstrecord, 0.5, type = 1),
           days90 = quantile(firstrecord, 0.90, type = 1),
           days95 = quantile(firstrecord, 0.95, type = 1),
           days99 = quantile(firstrecord, 0.99, type = 1),
           .groups = "drop"
           )

```


```{r}
summary_providers_firstrecord %>%
  reactable::reactable(filterable = TRUE)
```


```{r}
summary_providers_firstrecord %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(stat_name, " - ", name),
         loading = case_when(audit_date %in% c("20230819","20230824","20230829","20230901") ~ "Disrupted",
                                     TRUE ~ "Undisrupted")) %>%
  ggplot(aes(x = factor(category, 
                        levels = paste0(
                          c(
                            rep("Admissions per Provider", 4), 
                            rep("Discharged per Provider", 4), 
                            rep("Diags present per Provider", 4), 
                            rep("Arrivals per Provider", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        labels = paste0(
                          c(
                            rep("Admisions", 4), 
                            rep("Discharged", 4), 
                            rep("Discharged + diagcodes", 4), 
                            rep("ED Arrivals", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        ), 
             y = count, fill = source, color = loading)) +
		labs(
			x = "Percentile of providers sent first record",
			y = "No. of days after adm/arr date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  scale_color_manual(values = c("orange", "black")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 5)

```

#### Last records

```{r}
summary_providers_lastrecord <-
  summary_providers %>%
    group_by(source, stat_name, audit_date) %>%
    summarise(total = n_distinct(provider),
           days50 = quantile(lastrecord, 0.5, type = 1),
           days90 = quantile(lastrecord, 0.90, type = 1),
           days95 = quantile(lastrecord, 0.95, type = 1),
           days99 = quantile(lastrecord, 0.99, type = 1),
           .groups = "drop"
           )

```


```{r}
summary_providers_lastrecord %>%
  reactable::reactable(filterable = TRUE)
```


```{r}
summary_providers_lastrecord %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(stat_name, " - ", name),
         loading = case_when(audit_date %in% c("20230819","20230824","20230829","20230901") ~ "Disrupted",
                                     TRUE ~ "Undisrupted")) %>%
  ggplot(aes(x = factor(category, 
                        levels = paste0(
                          c(
                            rep("Admissions per Provider", 4), 
                            rep("Discharged per Provider", 4), 
                            rep("Diags present per Provider", 4), 
                            rep("Arrivals per Provider", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        labels = paste0(
                          c(
                            rep("Admisions", 4), 
                            rep("Discharged", 4), 
                            rep("Discharged + diagcodes", 4), 
                            rep("ED Arrivals", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        ), 
             y = count, fill = source, color = loading)) +
		labs(
			x = "Percentile of providers sent last record",
			y = "No. of days after adm/arr date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  scale_color_manual(values = c("orange", "black")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 8)

```

#### First and last records


```{r}
summary_providers_firstlast_record <-
  rbind(tibble(summary_providers_firstrecord, record_position = "first"),
        tibble(summary_providers_lastrecord, record_position = "last"))
```


```{r}
summary_providers_firstlast_record %>%
  filter(stat_name %in% c("Admissions per Provider", "Arrivals per Provider")) %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(record_position, " - ", stat_name, " - ", name),
         loading = case_when(audit_date %in% c("20230819","20230824","20230829","20230901") ~ "Disrupted",
                                     TRUE ~ "Undisrupted")) %>%
  ggplot(aes(x = factor(category, 
                        levels = paste0(
                          c(
                            rep("first - ", 4),
                            rep("last - ", 4)
                            ),
                          c(
                            rep("Admissions per Provider", 8), 
                            rep("Arrivals per Provider", 8)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        labels = paste0(
                          c(
                            rep("First ", 4),
                            rep("Last ", 4)
                            ),
                          c(rep("Admission record", 8), 
                            rep("Arrival record", 8)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        ), 
             y = count, fill = source, color = loading)) +
		labs(
			x = "Percentile of providers submitting records",
			y = "No. of days after adm/arr date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  scale_color_manual(values = c("orange", "black")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 8)


```


#### Records submitted without a disdate/diagcodes {.tabset}

This explains the similarity between APC subtypes

```{r}
summary_apc_subtype_differences <-
  monitor_results %>%
    filter(monitor_point_name == "sus_dates") %>%
    filter(substr(monitor_point_subname, 1, 13) %in% c( "apc_adm_date_")) %>%
    mutate(audit_date = substr(monitor_point_subname, regexpr(pattern = "date_", monitor_point_subname) + 5, 22)) %>%
    filter(stat_name %in% c("Admissions per Provider", "Discharged per Provider", "Diags present per Provider")) %>%
    filter(!is.na(value)) %>%
    mutate("num_value" = as.numeric(value)) %>%
    select(audit_date, run_date, stat_name, provider = groupby_value, num_value) %>%
    pivot_wider(names_from = stat_name, values_from = num_value) %>%
    mutate(no_disdate = `Admissions per Provider` - `Discharged per Provider`,
           disdate_no_diags = `Discharged per Provider` - `Diags present per Provider`,
           adm_no_diags = `Admissions per Provider` - `Diags present per Provider`)

```

##### Summary (first monitor date)

```{r}
summary_apc_subtype_differences %>%
  group_by(provider, audit_date) %>%
  arrange(run_date) %>%
  slice(1) %>%
  group_by(audit_date) %>%
  summarise(n_providers = n(),
            n_all_disdates_present = sum(no_disdate == 0),
            p_all_disdates_present = round(n_all_disdates_present/n_providers*100),
            n_all_diags_present_when_disdate_present = sum(disdate_no_diags == 0),
            n_all_diags_present = sum(adm_no_diags == 0),
            p_all_diags_present = round(n_all_diags_present/n_providers*100)
            ) %>%
  reactable::reactable()
  
```


```{r}
summary_providers_firstrecord %>%
  filter(source != "ECDS") %>%
  pivot_longer(cols = starts_with("days"),
               names_prefix = "days",
               values_to = "count") %>%
  mutate(category = paste0(stat_name, " - ", name),
         loading = case_when(audit_date %in% c("20230819","20230824","20230829","20230901") ~ "Disrupted",
                                     TRUE ~ "Undisrupted")) %>%
  ggplot(aes(x = factor(category, 
                        levels = paste0(
                          c(
                            rep("Admissions per Provider", 4), 
                            rep("Discharged per Provider", 4), 
                            rep("Diags present per Provider", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        labels = paste0(
                          c(
                            rep("Admisions", 4), 
                            rep("Discharged", 4), 
                            rep("Discharged + diagcodes", 4)
                            ), 
                          rep(c(" - 50", " - 90", " - 95", " - 99"), 4)),
                        ), 
             y = count, fill = source, color = loading)) +
		labs(
			x = "Percentile of providers sent first record",
			y = "No. of days after adm/arr date"
			) +
  scale_y_continuous(breaks = seq(0, 510, by = 30)) +
  scale_color_manual(values = c("orange", "black")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 5)

```

